The Hot Hand in Call of Duty Search and Destroy

### 2022-08-21 paper/paper.tex
feedback thru old p6

\title{They're Icing up! Team Momentum in Round-Based Esports}

\author{
    Tony ElHabr
   \\
     \\
   \\
  \texttt{\href{mailto:anthonyelhabr@gmail.com}{\nolinkurl{anthonyelhabr@gmail.com}}} \\
  }

@@

\begin{abstract}
My research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD). First, I find evidence
supprting the hypothesis that round win probability can be modeled as a
constant across the series, although not at the naive 50\%. Second, I
examine post-streak round win probability, given the series outcome. I
find no evidence that post-streak win rate is significantly different
than expected for combinations of streak length, series length, and
series winner, suggesting that the perception of momentum at the team
level is just that, perception.
\end{abstract}

@@

\hypertarget{game-description}{%
\subsection{Game Description}\label{game-description}}

@@

In professional CoD SnD, a team take turns playing offense and defense
every round. They must win six rounds to win the series.\footnote{A
  maximum of 11 even rounds can be played. There is no ``sudden death''
  or ``win by two'' rule like there are for SnD equivalent in
  professional Counter-Strike and Valorant matches.} A round can end in
one of five ways:

@@

I adopt the terminology ``series'' to refer to what CoD SnD players
typically call a ``match'', so as to emulate the terminology of playoff
series in professional leagues like the National Basketball Association,
National Hockey League, and Major League Baseball. A ``game'' or a
``match'' in such leagues is analogous to a ``round'' of CoD SnD.
Further, it is analogous to a trial in statistical analysis of Bernoulli
events.

@@

The empirical offensive round win percentage across all rounds is
\(\tau_O\) = 47.8\%.\footnote{Offensive round win percentage has been
  nearly constant across the three games during the CDL era: 1. 47.2\%
  in MW (2020) 2. 47.9\% in Cold War (2021) 3. 48.1\% in Vanguard (2022)}
Table \ref{tbl:cod-o-win-prop-by-series-state} shows round win
percentages by series ``state'' (i.e.~the number of round wins by each
team prior to an upcoming round). Offensive round win rate is not quite
constant, although never veers more than 10\% from this global
average.\footnote{At tournaments, higher-seeded (``better'') teams
  choose whether they want to start on offense or defense in either the
  SnD series played as the second game or the fifth game in the
  best-of-five format used in tournaments, and their opponent is
  assigned to play offense in the first round of the SnD slot (second or
  fifth game) not chosen. ({``Competitive Settings \textbar{} Call of
  Duty League,''} n.d.) (Other game variants like Hardpoint are played
  as the first, third, and fourth games in the matchup.) Anecdotally,
  the higher-seeded teams tend to choose to play defense in the first
  round of the SnD series played as the second game, although this
  choice is not universal. With online qualifiers, teams all play each
  other twice throughout the season. Each team gets to play the
  higher-seeded role once in their matchups, despite their win-loss
  records.}

\begin{longtable}{crrrrrr}
\caption{Offensive round win rates for the upcoming round, given both the offensive and defensive team's prior number of round wins}\label{tbl:cod-o-win-prop-by-series-state} \\
\toprule
& \multicolumn{6}{c}{Offense round wins} \\ 
\cmidrule(lr){2-7}
Defense round wins & 0 & 1 & 2 & 3 & 4 & 5 \\ 
\midrule
0 & 47.8\%
(852) & 46.6\%
(408) & 43.1\%
(216) & 43.5\%
(115) & 43.3\%
(67) & 40.5\%
(37) \\ 
1 & 48.6\%
(444) & 49.3\%
(418) & 51.5\%
(309) & 43.4\%
(205) & 43.3\%
(120) & 39.4\%
(99) \\ 
2 & 52.8\%
(218) & 48.9\%
(305) & 48.9\%
(315) & 46.6\%
(262) & 48.7\%
(189) & 42.1\%
(133) \\ 
3 & 54.5\%
(123) & 46.0\%
(200) & 49.6\%
(250) & 45.6\%
(248) & 44.4\%
(214) & 44.8\%
(174) \\ 
4 & 56.9\%
(65) & 54.5\%
(145) & 47.2\%
(193) & 44.7\%
(228) & 55.2\%
(221) & 50.5\%
(208) \\ 
5 & 47.4\%
(38) & 49.4\%
(83) & 47.1\%
(136) & 50.9\%
(175) & 45.2\%
(177) & 46.0\%
(202) \\ 
\bottomrule
\end{longtable}

Anecdotally, observers swear by the existence of momentum in CoD SnD
series, to the extent that vernacular has been developed to describe
such phenomenon. People say that teams ``ice up'' when the team wins
several consecutive rounds to come back to win a series, often
culminating with a series-clinching round win.\footnote{This vernacular
  is not specific to CoD SnD in specific, and is not only used in the
  contexts of streaks.} Further, viewers have come to embrace the
``5-3'' phenomenon, where teams win three consecutive rounds after
facing a 5-3 deficit to win 6-5. There is even a term for the rare 0-5
comeback---a ``full sail''.

@@

There have been a handful of studies of the distribution of games played
in a series of a professional sport. Most assume a constant probability
\(\phi\) of a given team winning a game in the series, regardless of the
series state. Mosteller (1952) observed that the American League had
dominated the National League in Major League Baseball's (MLB) World
Series matchups, implying that games should not be modeled with a
constant (null) \(\phi_0 = 0.5\). Mosteller proposed three approaches
for identifying the optimal constant probability value of the stronger
team in the World Series, finding \(\phi \approx 0.65\) in each case:
(1) solving for \(p\) from the empirical average number of games won by
the loser of the series, which he called the ``method of moments''
approach; (2) maximizing the likelihood that the sample would have been
drawn from a population in which the probability of a team winning a
game is constant across the series (i.e.~maximum likelihood), and (3)
minimizing the chi-square goodness of fit statistic \(\chi^2\) for
\(\phi\).

@@

Studying both player streaks and team streaks in basketball, in both
observational and controlled settings. Gilovich et al. (1985),
henceforth GVT, do not find evidence for the hot hand phenomenon.
However, Miller and Sanjurjo (2018), henceforth MS, refuted the
conclusions of GVT, proving a framework for quantifying streak selection
bias, which effectively works in the manner posited by the gambler's
fallacy\footnote{Streak selection bias, ``in conjunction with a
  quasi-Bayesian model of decision making under sample size
  neglect\ldots{} provides a novel structural candidate explanation for
  the persistence of gambler's fallacy beliefs'', per MS.}.
Specifically, MS say that a ``bias exists in a common measure of the
conditional dependence of present outcomes on streaks of past outcomes
in sequential data'' implying that, under i.i.d. conditions, ``the
proportion of successes among the outcomes that immediately follow a
streak of consecutive successes is expected to be strictly less than the
underlying (conditional) probability of success''. When applying streak
selection bias to GVT.'s data, MS come to the opposite conclusions as
GVT.

@@

\hypertarget{my-contribution}{%
\subsection{My contribution}\label{my-contribution}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Both have economic aspects that can create clear advantages for one
  side in a given round, given how prior rounds played out. CoD has no
  such equivalent, except for perhaps ``kill streaks'', which
  infrequently play a role.
\item
  Teams play either offense or defense for many consecutive rounds. (The
  former has a 14-14-1 format and the latter uses a 12-12-1 format.) On
  the other hand, teams in CoD SnD rotate roles every round, analogous
  to a 1-1-1-1-1-1-1 format for home advantage in best-of-seven series
  for professional sports like the MLB, NBA, and NHL.\footnote{1-1-1-1-1-1-1
    is not used today in these leagues, but it was at least one in each
    league.} While theoretically one might be able to account for any
  kind of format, such as a 5-5-1, the rotation of team sides every
  round is convenient for convincing ourselves that rounds could
  reasonably be modeled as i.i.d. Bernoulli trials.
\item
  Both Counter-Strike and Valorant have overtime rules---team must win
  by two rounds---which can make end-of-series sequences difficult to
  model. CoD SnD does not have overtime rules.
\end{enumerate}

To my knowledge, there is no existing public statistical research on CoD
SnD format, beyond descriptive analysis on social media. While
intra-round trends may be more directly applicable to teams looking for
an advantage on their competition, broader investigation on a common
concept like the hot-hand phenomenon in a new field---particularly one
that is less subjected to factors that may be difficult to control for,
e.g.~weather---should hopefully be useful as a precedent or supplement
for future researchers.

@@

The formula for the expected proportion of series a best-of-\(s\) format
ending in \(r\) rounds (\(r \leq s\)), \(\hat{\Phi}(r)\), for a constant
round win probability \(\phi\) (\(0 \leq \phi \leq 1\)) for one
team\footnote{If \(\phi > 0.5\), then one might say that this team is
  the better team (known in hindsight).} is

\begin{equation}\protect\hypertarget{eq:m}{}{
m = \frac{s + 1}{2}
}\label{eq:m}\end{equation}

\begin{equation}\protect\hypertarget{eq:series-length}{}{
\hat{\Phi}(r) = \frac{(r - 1)!}{(m - 1)!(s - r)!}(\phi^{m}(1 - \phi)^{r - m} + \phi^{r - m}(1 - \phi)^m)
}\label{eq:series-length}\end{equation}

\(s = 11\) rounds for CoD SnD.\footnote{\(\phi\) and other symbols are
  selected so as to reserve symbols like \(p\) for usage in other
  contexts without causing confusion for the reader. Upper-case symbol,
  e.g.~\(\Phi\), are consistently used in this paper to represent
  proportions, while lower-case symbols corresponding to proportions
  represent probabilities. ``Hats'', e.g.~\(\hat{\Phi}\), are used to
  convey expectations, while ``bare'' symbols mirroring hat-ed symbols
  convey observational data.}

To evaluate the constant round win probability null hypothesis---that
is, that the expected and observed round win rates, \(\hat{\Phi}(r)\)
and \(\Phi(r)\) respectively, are equal to one another---we can compute
the chi-square goodness of fit statistic, \(\chi^2\),

\begin{equation}\protect\hypertarget{eq:chi-squ}{}{
\chi^2 = \sum^R \frac{(\Phi(r) - \hat{\Phi}(r))^2}{\hat{\Phi}(r)}.
}\label{eq:chi-squ}\end{equation}

\(r \in R = [6, 7, 8, 9, 10, 11]\) for CoD SnD.

@@

\begin{equation}\protect\hypertarget{eq:pwr}{}{
p_0(\text{win} | r) = p^{+|r}_0 = \{
\begin{array}{cl}
\frac{m}{r}, & \text{team wins series} \\
\frac{r - m}{r}, & \text{team loses series}
\end{array}
}
\label{eq:pwr}
\end{equation}

@@

As shown by MS with their Theorem 1, we should expect the proportion of
rounds wins immediately following a streak of \(k\) rounds wins for a
series lasting \(r\) rounds, \(\hat{P}^{+|kr}_{MS}\), to be strictly
less than \(p^{+|r}\).\footnote{See Appendix E.1 of MS for the proof.}
Unfortunately, there does not exist a closed form representation for the
expected value of \(\hat{P}^{+|kr}_{MS}\) for \(k > 1\). Nonetheless,
one may recursively build a collection of dictionaries to compute the
expectation analytically, or, more conveniently, one may run simulations
to estimate the expected value, as I choose to do.\footnote{I have
  adapted the code from Vafa (2017), which implements MS's framework
  (2018).}

Although it would be convenient to emulate the hypothesis testing for
difference of proportions performed by MS (and GVT), my context is
fundamentally different from that of MS, who focus on longitudinal data
in controlled settings.\footnote{GVT also performs statistical tests on
  shots from players in live games, i.e.~``observational'' data, but
  they note that their findings are likely affected player shot
  selection in the face of defensive strategy by the opposing team.} The
number of trials is fixed in their experimental designs, but in CoD SnD,
the number of rounds played is determined as a function of the max
number of possible rounds (\(s\)) and whether or not the team wins the
series. The Bernoulli trial success probability, i.e.~a round win in CoD
SnD, is not independent of the opponent. Consequently, a statistical
test of the difference in \(\hat{P}^{+|kr}\) and \(\hat{P}^{-|kr}\), as
performed by MS to evaluate their hypothesis regarding post-streak
success rates, is not completely appropriate, although useful as a
reference.

@@

One might consider another form of the expected proportion of rounds won
immediately after a streak of \(k\) round wins in a best-of-\(s\) series
given the length of the series (\(r\) rounds), the ``notional''
proportion \(\hat{P}^{+|kr}_0\). The Bernoulli round win probability
underlying \(\hat{P}^{+|kr}_0\) is

\begin{equation}\protect\hypertarget{eq:pwkr}{}{
p_0(\text{win} | k, r) = p^{+|kr}_0 =\{
\begin{array}{cl}
\frac{m - k}{r - k}, & \text{team wins series} \\
\frac{s - m - k}{r - k}, & \text{team loses series}
\end{array}
}
\label{eq:pwkr}
\end{equation}

@@

\begin{equation}\protect\hypertarget{eq:binom}{}{
\Pr(r^w) = {\binom {r}{r^+}} \omega^{r^+}(1-\omega)^{r-r^+}.
}\label{eq:binom}\end{equation}

Plugging in the notional proportion \(\hat{p}^{+|kr}_0\) for the null
\(\omega_0\) and the observed proportion \(P^{+|kr}\) for \(\omega\)
(treating proportions as probabilities), we can evaluate the null
hypothesis that the observed probability is equal to the notional
probability. If we can reject this null hypothesis, then we might
consider team momentum, represented by post-streak success, plausible.

One can perform the same binomial test for the MS's
streak-selection-adjusted proportion, \(\hat{P}^{+|kr}_{MS}\). FOOTY:
Given the caveats mentioned before, the results should be heeded with
caution.

@@

\begin{equation}\protect\hypertarget{eq:pwkri}{}{
p_0(\text{win} | k, r, i) = p^{+|kri}_0 =\{
\begin{array}{cl}
\frac{m - k}{r - k}, & \text{team wins series}, i \neq r \\
1, & \text{team wins series}, i = r \\
\frac{s - m - k}{r - k}, & \text{team loses series}, i \neq r \\
0, & \text{team loses series}, i = r
\end{array}
}
\label{eq:pwkri}
\end{equation}

Again, one might apply a binomial test to evaluate the hypothesis that
the expected proportions, both \(\hat{P}^{+|kri}_0\) and
\(\hat{P}^{+|kri}_{MS}\), are equal to the observed proportion
\(P^{+|kr}\).

@@

To incorporate my findings regarding constant round win probability, I
can specify that \(\Pr(r^+) = \phi\) (and, conversely, that
\(\Pr(r^-) = 1 - \phi\)).

@@

First, I investigate the constant probability assumption and the
distribution of rounds played in a series. Chance's (2020) work is
closely related to mine, and, in fact, provides a guide for this
investigation. Afterwards, I investigate post-streak win rates and the
context of momentum.

@@

Using Equation \ref{eq:chi-squ}, I find that \(\chi^2 = 16.0\)
(\(p\)-value of 0.0068) for the naive constant round win probability
\(\phi_0 = 0.5\). Thus, I can comfortably reject the constant
probability hypothesis for the null \(\phi_0 = 0.5\), even at a
confidence level of \(\alpha = 0.01\).

@@

\begin{longtable}{rrrr}
\caption{The expected proportion of series lasting $r$ rounds ($\hat{\Phi}_0(r)$) in a best-of-11 format, wehre $r \in R = [6, 7, 8, 9, 10, 11]$ under the assumption that each team has a constant round win probability $\phi_0 = 0.5$. The observed frequencies for CoD SnD are shown as a count $N(r)$ and as a proportion $\Phi(r)$ of all series ($\sum^R N(r)$).}\label{tbl:cod-prob-series-lasting-r-rounds} \\
\toprule
$r$ & $\hat{\Phi}_0(r)$ & $\Phi(r)$ & $N(r)$ \\ 
\midrule
6 & $3.1\%$ & $4.7\%$ & 40 \\ 
7 & $9.4\%$ & $11.9\%$ & 101 \\ 
8 & $16.4\%$ & $16.5\%$ & 141 \\ 
9 & $21.9\%$ & $21.7\%$ & 185 \\ 
10 & $24.6\%$ & $21.5\%$ & 183 \\ 
11 & $24.6\%$ & $23.7\%$ & 202 \\ 
\bottomrule
\end{longtable}

Table \ref{tbl:mosteller-methods-results} shows the alternate values for
the constant round win probability that I find when applying the three
methods suggested by Mosteller (1952). Each is approximately or equal to
0.575, and, when applying Equation \ref{eq:chi-squ}, each results in a
\(\chi^2\) value for which I cannot reject the constant probability null
hypothesis.

\begin{longtable}[]{@{}lrr@{}}
\caption{Alternate estimates of the constant probability ($\phi$) for winning a given round in a CoD SnD, applying the three methods suggested by Mosteller (1952), in addition to the naive ($\phi_0 = 0.5$).}\label{tbl:mosteller-methods-results} \\
\toprule()
Method & $\phi$ & $\chi^2$ (p-value) \\
\midrule()
\endfirsthead
\toprule()
Method & $\phi(r)$ & $\chi^2$ (p-value) \\
\midrule()
\endhead
0. Naive & 0.5000 & 16.0 (\textless=0.01) \\
1. Method of moments & 0.5725 & 3.6 (0.6) \\
2. Maximum likelihood & 0.5750 & 3.5 (0.62) \\
3. Minimum ($\chi^2$) & 0.5775 & 3.5 (0.62) \\
\bottomrule()
\end{longtable}

>>>>

\title{The Hot Hand in Call of Duty Search and Destroy}

\author{
    Tony ElHabr
   \\
     \\
   \\
  \texttt{\href{mailto:anthonyelhabr@gmail.com}{\nolinkurl{anthonyelhabr@gmail.com}}} \\
   \And
    Andrew ElHabr
   \\
    School of Industrial and Systems Engineering (ISyE) \\
    Georgia Institute of Technology \\
  Atlanta, Georgia, USA \\
  \texttt{\href{mailto:andrewelhabr@gatech.edu}{\nolinkurl{andrewelhabr@gatech.edu}}} \\
  }

@@

\begin{abstract}
Our research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD). First, we find evidence
supprting the hypothesis that round win probability can be modeled as a
constant across the series, although not at the naive 50\%. Second, we
examine post-streak round win probability, given the series outcome. We
find no evidence that post-streak win rate is significantly different
than expected for combinations of streak length, series length, and
series winner, suggesting that the perception of momentum at the team
level is just that, perception.
\end{abstract}

\keywords{
    esports
   \and
    Call of Duty
   \and
    hot hand
   \and
    runs test
  }

@@

\hypertarget{call-of-duty-search-and-destroy}{%
\subsection{Call of Duty Search and
Destroy}\label{call-of-duty-search-and-destroy}}

@@

In professional CoD SnD, the two teams take turns playing offense and
defense every round. They must win six rounds to win the
series.\footnote{A maximum of 11 even rounds can be played. There is no
  ``sudden death'' or ``win by two'' rule like there are for SnD
  equivalent in professional Counter-Strike and Valorant matches.} A
round can end in one of five ways:

@@

We adopt the terminology ``series'' to refer to what CoD SnD players
typically call a ``match'', so as to emulate the terminology of playoff
series in professional leagues like the National Basketball Association,
National Hockey League, and Major League Baseball. A ``game'' or a
``match'' in such leagues is analogous to a ``round'' of CoD SnD.
Further, it is analogous to a trial in statistical analysis of Bernoulli
events.

SnD is not the only game type played in competitive CoD. Teams play in a
head-to-head, best-of-five (first-to-three wins) format, where SnD is
always played as the second and fifth game mode (if a fifth variant is
needed). The best-of-five across game types could also be called a
``series'', but since we analyze only the SnD game types, we refer to
SnD games as series (of rounds).

Anecdotally, most observers swear by the existence of momentum in CoD
SnD series, to the extent that vernacular has been developed to describe
such phenomenon. People say that teams ``ice up'' when the team wins
several consecutive rounds to come back to win a series, often
culminating with a series-clinching round win.\footnote{This vernacular
  is not specific to CoD SnD in specific, and is not only used in the
  contexts of streaks.} Further, viewers have come to embrace the
``5-3'' phenomenon, where teams win three consecutive rounds after
facing a 5-3 deficit to win 6-5. There is even a term for the rare 0-5
comeback---a ``full sail''.

@@

The observed offensive round win percentage across all rounds is
\(\tau_O\) = 47.8\%.\footnote{Offensive round win percentage has been
  nearly constant across the three games during the CDL era: 1. 47.2\%
  in MW (2020) 2. 47.9\% in Cold War (2021) 3. 48.1\% in Vanguard (2022)}
Table \ref{tbl:cod-o-win-prop-by-series-state} shows round win
percentages by series ``state'' (i.e.~the number of round wins by each
team prior to an upcoming round). Offensive round win rate is not quite
constant, although never veers more than 10\% from this overall average.

\begin{longtable}{crrrrrr}
\caption{Offensive round win rates for the upcoming round, given both the offensive and defensive team's prior number of round wins. Numbers in parentheses are sample sizes.}\label{tbl:cod-o-win-prop-by-series-state} \\
\toprule
& \multicolumn{6}{c}{Offense round wins} \\ 
\cmidrule(lr){2-7}
Defense round wins & 0 & 1 & 2 & 3 & 4 & 5 \\ 
\midrule
0 & 47.8\%
(852) & 46.6\%
(408) & 43.1\%
(216) & 43.5\%
(115) & 43.3\%
(67) & 40.5\%
(37) \\ 
1 & 48.6\%
(444) & 49.3\%
(418) & 51.5\%
(309) & 43.4\%
(205) & 43.3\%
(120) & 39.4\%
(99) \\ 
2 & 52.8\%
(218) & 48.9\%
(305) & 48.9\%
(315) & 46.6\%
(262) & 48.7\%
(189) & 42.1\%
(133) \\ 
3 & 54.5\%
(123) & 46.0\%
(200) & 49.6\%
(250) & 45.6\%
(248) & 44.4\%
(214) & 44.8\%
(174) \\ 
4 & 56.9\%
(65) & 54.5\%
(145) & 47.2\%
(193) & 44.7\%
(228) & 55.2\%
(221) & 50.5\%
(208) \\ 
5 & 47.4\%
(38) & 49.4\%
(83) & 47.1\%
(136) & 50.9\%
(175) & 45.2\%
(177) & 46.0\%
(202) \\ 
\bottomrule
\end{longtable}

On the extremes, 6-0 sweeps make up 4.7\% of all series, while 6-5
series make up 23.7 of series.

At tournaments, higher-seeded (``better'') teams choose whether they
want to start on offense in the SnD series played as the second game or
the fifth game, and their opponent is assigned to play offense in the
first round of the SnD slot not chosen.\footnote{``Hardpoint'' has been
  played as the first and fourth games in a matchup; the variant for the
  third game was Domination in the 2020 season, and Control in the 2021
  and 2022 seasons.}\footnote{Anecdotally, the higher-seeded teams tend
  to choose to play defense in the first round of the SnD series played
  as the second game, although this choice is not consistent across
  teams, or with the same team over time.} ({``Competitive Settings
\textbar{} Call of Duty League,''} n.d.) With online qualifiers, teams
all play each other twice throughout the season. Each team gets to play
the higher-seeded role once in their matchups, despite their win-loss
records. Overall, we might say that bias due to stronger teams playing
defense (slightly advantageous) more often is very minimal.

@@

There have been a handful of studies of the distribution of games played
in a series of a professional sport. Most assume a constant probability
\(\phi\) of a given team winning a game in the series, regardless of the
series state. Mosteller (1952) observed that the American League had
dominated the National League in Major League Baseball's (MLB) World
Series matchups, implying that games should not be modeled with a
constant (null) \(\phi_0 = 0.5\). Mosteller proposed three approaches
for identifying the optimal constant probability value of the stronger
team in the World Series, finding \(\phi \approx 0.65\) in each case:
(1) solving for \(\phi\) from the observed average number of games won
by the loser of the series, which he called the ``method of moments''
approach; (2) maximizing the likelihood that the sample would have been
drawn from a population in which the probability of a team winning a
game is constant across the series (i.e.~maximum likelihood), and (3)
minimizing the chi-square goodness of fit statistic \(\chi^2\) for
\(\phi\).

@@

Studying both player streaks and team streaks in basketball, in both
observational and controlled settings. Gilovich et al. (1985),
henceforth GVT, do not find evidence for the hot hand phenomenon.
However, Miller and Sanjurjo (2018), henceforth MS, provide a framework
for quantifying streak selection bias, which effectively works in the
manner posited by the gambler's fallacy. Specifically, MS say that a
``bias exists in a common measure of the conditional dependence of
present outcomes on streaks of past outcomes in sequential data''
implying that, under i.i.d. conditions, ``the proportion of successes
among the outcomes that immediately follow a streak of consecutive
successes is expected to be strictly less than the underlying
(conditional) probability of success''. When applying streak selection
bias to GVT's data, MS come to the opposite conclusions as GVT.

@@

\hypertarget{our-contribution}{%
\subsection{Our contribution}\label{our-contribution}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Both have economic aspects that can create clear advantages for one
  side in a given round, given how prior rounds played out. CoD has no
  such equivalent, except for perhaps ``kill streaks'', which
  infrequently play a role.
\item
  Teams play either offense or defense for many consecutive rounds. (The
  former has a 14-14-1 format and the latter uses a 12-12-1 format.) On
  the other hand, teams in CoD SnD rotate roles every round, analogous
  to a 1-1-1-1-1-1-1 format for home advantage in best-of-seven series
  for professional sports like the MLB, NBA, and NHL.\footnote{1-1-1-1-1-1-1
    is not used today in these leagues, but it was at least once in each
    league.} While theoretically one might be able to account for any
  kind of format, such as a 5-5-1, the rotation of team sides every
  round is convenient for convincing ourselves that rounds could
  reasonably be modeled as i.i.d. Bernoulli trials.
\item
  Both Counter-Strike and Valorant have overtime rules---team must win
  by two rounds---which can make end-of-series sequences difficult to
  model. CoD SnD does not have overtime rules.
\end{enumerate}

To our knowledge, there is no existing public statistical research on
CoD SnD format, beyond descriptive analysis on social media. While
intra-round trends may be more directly applicable to teams looking for
an advantage on their competition, broader investigation on a common
concept like the hot-hand phenomenon in a new field---particularly one
that is less subjected to factors that may be difficult to control for,
e.g.~weather---should be useful as a precedent or supplement for future
researchers.

@@

In a best-of-\(s\) format, assuming a constant round win probability
\(\phi\) (\(0 \leq \phi \leq 1\)) for one team\footnote{If
  \(\phi > 0.5\), we can say that this is the stronger team.}, the
probability that a series ends in \(r\) rounds (\(r \leq s\)),
\(\hat{\Phi}(r)\) is

\begin{equation}\label{eq:m}
m = \frac{s + 1}{2}
\end{equation}

\begin{equation}\label{eq:series-length}
\hat{\Phi}(r) = \frac{(r - 1)!}{(m - 1)!(s - r)!}(\phi^{m}(1 - \phi)^{r - m} + \phi^{r - m}(1 - \phi)^m)
\end{equation}

\(s = 11\) rounds for CoD SnD.\footnote{\(\phi\) and other symbols are
  selected so as to reserve symbols like \(p\) for usage in other
  contexts without causing confusion for the reader. Upper-case symbol,
  e.g.~\(\Phi\), are consistently used in this paper to represent
  proportions, while lower-case symbols are used to represent Bernoulli
  trials success probability. ``Hats'', e.g.~\(\hat{\Phi}\), are used to
  convey expectations, while bare symbols symbols convey observational
  data.} For example, assuming \(\phi = 0.5\), the probability of a
series ending in 9 rounds in CoD is

\[
\hat{\Phi}(9) = \frac{(9 - 1)!}{(6 - 1)!(11 - 9)!}(0.5^{6}(1 - 0.5)^{9 - 6} + 0.5^{9 - 6}(1 - 0.5)^6) = 56 (0.5^9 + 0.5^9) = 0.21875.
\]

To evaluate the constant round win probability null hypothesis---that
is, that the expected and observed round win rates, \(\hat{\Phi}(r)\)
and \(\Phi(r)\) respectively, are equal to one another---we can compute
the chi-square goodness of fit statistic

\begin{equation}\label{eq:chi-squ}
\chi^2 = \sum^{r \in R} \frac{(\Phi(r) - \hat{\Phi}(r))^2}{\hat{\Phi}(r)}.
\end{equation}

in which \(R = [6, 7, 8, 9, 10, 11]\) for CoD SnD.

@@

\begin{equation}\label{eq:pwr}
p_0(\text{win} | r) = p^{+|r}_0 = \begin{cases} 
\frac{m}{r} & \text{team wins series} \\ 
\frac{r - m}{r} & \text{team loses series}  
\end{cases}
\end{equation}

@@

As shown by MS with their Theorem 1, we should expect the proportion of
rounds wins immediately following a streak of \(k\) rounds wins for a
series lasting \(r\) rounds, \(\hat{P}^{+|k,r}_{MS}\), to be strictly
less than \(p^{+|r}\).\footnote{See Appendix E.1 of MS for the proof.}
Unfortunately, there does not exist a closed form representation for the
expected value of \(\hat{P}^{+|k,r}_{MS}\) for \(k > 1\). Nonetheless,
one may recursively build a collection of dictionaries to compute the
expectation analytically, or, more conveniently, one may run simulations
to estimate the expected value, as We choose to do.\footnote{We have
  adapted the code from Vafa (2017), which implements MS's framework
  (2018).}

Although it would be convenient to emulate the hypothesis testing for
difference of proportions performed by MS (and GVT), our context is
fundamentally different from that of MS, who focus on longitudinal data
in controlled settings.\footnote{GVT also performs statistical tests on
  shots from players in live games, i.e.~``observational'' data, but
  they note that their findings are likely affected player shot
  selection in the face of defensive strategy by the opposing team.} The
number of trials is fixed in their experimental designs, but in CoD SnD,
the number of rounds played is determined as a function of the max
number of possible rounds (\(s\)) and whether or not the team wins the
series. The Bernoulli trial success probability, i.e.~a round win in CoD
SnD, is not independent of the opponent. Consequently, a statistical
test of the difference in \(\hat{P}^{+|k,r}\) and \(\hat{P}^{-|kr}\), as
performed by MS to evaluate their hypothesis regarding post-streak
success rates, is not completely appropriate, although useful as a
reference.

@@

One can consider another form of the expected proportion of rounds won
immediately after a streak of \(k\) round wins in a best-of-\(s\) series
given the length of the series (\(r\) rounds), the ``notional''
proportion \(\hat{P}^{+|k,r}_0\). The Bernoulli round win probability
underlying \(\hat{P}^{+|k,r}_0\) is

\begin{equation}\label{eq:pwkr}
p_0(\text{win} | k, r) = p^{+|k,r}_0 = \begin{cases}
\frac{m - k}{r - k} & \text{team wins series} \\
\frac{s - m - k}{r - k} & \text{team loses series}
\end{cases}
\end{equation}

@@

\begin{equation}\label{eq:binom}
\Pr(r^w) = {\binom {r}{r^+}} \omega^{r^+}(1-\omega)^{r-r^+}.
\end{equation}

Treating the notional proportion \(p^{+|k,r}_0\) as the null
\(\omega_0\) and plugging in the observed proportion \(P^{+|k,r}\) for
\(\omega\) in Equation \ref{eq:binom} (treating proportions as
probabilities), we can evaluate the null hypothesis that the observed
probability is equal to the notional probability. If we can reject this
null hypothesis, then we can consider team momentum, represented by
post-streak success, plausible.

One can perform the same binomial test for the MS's
streak-selection-adjusted proportion, \(\hat{P}^{+|k,r}_{MS}\). FOOTY:
Given the caveats mentioned before, the results should be heeded with
caution.

@@

\begin{equation}\protect\hypertarget{eq:pwkri}{}{
p_0(\text{win} | k, r, i) = p^{+|k,r,i}_0 =\{
\begin{array}{cl}
\frac{m - k}{r - k}, & \text{team wins series}, i \neq r \\
1, & \text{team wins series}, i = r \\
\frac{s - m - k}{r - k}, & \text{team loses series}, i \neq r \\
0, & \text{team loses series}, i = r
\end{array}
}
\label{eq:pwkri}
\end{equation}

Again, one might apply a binomial test to evaluate the hypothesis that
the expected proportions, both \(\hat{P}^{+|k,r,i}_0\) and
\(\hat{P}^{+|k,r,i}_{MS}\), are equal to the observed proportion
\(P^{+|k,r}\).

@@

To incorporate our findings regarding constant round win probability, We
can specify that \(\Pr(r^+) = \phi\) (and, conversely, that
\(\Pr(r^-) = 1 - \phi\)).

@@

First, We investigate the constant probability assumption and the
distribution of rounds played in a series. Chance's (2020) work is
closely related to mine, and, in fact, provides a guide for this
investigation. Afterwards, We investigate post-streak win rates within
the context of momentum.

@@

Using Equation \ref{eq:chi-squ}, We find that \(\chi^2 = 16.0\)
(\(p\)-value of 0.0068) for the naive constant round win probability
\(\phi_0 = 0.5\). Thus, We can comfortably reject the constant
probability hypothesis for the null \(\phi_0 = 0.5\), even at a
confidence level of \(\alpha = 0.01\).

@@

\begin{longtable}{rrrr}
\caption{The expected proportion of series lasting $r$ rounds ($\hat{\Phi}_0(r)$) in a best-of-11 format, where $r \in R = [6, 7, 8, 9, 10, 11]$ under the assumption that each team has a constant round win probability $\phi_0 = 0.5$. The observed frequencies for CoD SnD are shown as a count $N(r)$ and as a proportion $\Phi(r)$ of all series ($\sum^{r \in R} N(r)$).}\label{tbl:cod-prob-series-lasting-r-rounds} \\
\toprule
$r$ & $N(r)$ & $\Phi(r)$ & $\hat{\Phi}_0(r)$ \\ 
\midrule
6 & 40 & $4.7\%$ & $3.1\%$ \\ 
7 & 101 & $11.9\%$ & $9.4\%$ \\ 
8 & 141 & $16.5\%$ & $16.4\%$ \\ 
9 & 185 & $21.7\%$ & $21.9\%$ \\ 
10 & 183 & $21.5\%$ & $24.6\%$ \\ 
11 & 202 & $23.7\%$ & $24.6\%$ \\ 
\bottomrule
\end{longtable}

Table \ref{tbl:mosteller-methods-results} shows the alternate values for
the constant round win probability that We find when applying the three
methods suggested by Mosteller (1952). Each is approximately or equal to
0.575, and, when applying Equation \ref{eq:chi-squ}, each results in a
\(\chi^2\) value for which We cannot reject the constant probability
null hypothesis.

\begin{longtable}[]{@{}lrr@{}}
\caption{Alternate estimates of the constant probability ($\phi$) for winning a given round in a CoD SnD, applying the three methods suggested by Mosteller (1952), in addition to the naive ($\phi_0 = 0.5$).}\label{tbl:mosteller-methods-results} \\
\toprule()
Method & $\phi$ & $\chi^2$ (p-value) \\
\midrule()
\endfirsthead
\toprule()
Method & $\phi(r)$ & $\chi^2$ (p-value) \\
\midrule()
\endhead
0. Naive & 0.5000 & 16.0 ($\leq$ 0.01) \\
1. Method of moments & 0.5725 & 3.6 (0.6) \\
2. Maximum likelihood & 0.5750 & 3.5 (0.62) \\
3. Minimum ($\chi^2$) & 0.5775 & 3.5 (0.62) \\
\bottomrule()
\end{longtable}
